The Development of a Collaborative Tool to Teach Debugging

### 2020-12-06 paper/paper.tex
Final updates to paper

\begin{abstract}
  \linespread{1}
  Debugging is rarely formally taught, despite being used by
  programmers every day.  Research indicates that it is valuable to
  teach debugging, and suggests that teaching it collaboratively may
  be maximally effective.  The goal of this project is to create a
  collaborative debugger.  The debugger aims to be the ideal platform
  to teach and learn debugging.  This paper briefly reviews relevant
  literature which covers teaching debugging and that which focuses on
  teaching programming collaboratively.  Most of the paper is devoted
  to the design of the collaborative debugger.
\end{abstract}

@@

\begin{itemize}
\item Keith O'Hara, my senior project advisor, for his constant
  thoughtfulness about my project, his advice about the future, and
  for giving me the opportunity to tutor for his classes.
\item Mar Hicks and Kyle Hale, professors whose classes at IIT changed
  my perspective on computer science.
\item All my professors at Bard, particularly Olga Voronina, Wakako
  Suzuki, and Nate Shockey, whose encouragement of my interests
  outside of computer science has been instrumental in continuing my
  passion for it.
\item My family and friends, for their constant love and support.
\end{itemize}

@@

Debugging is invaluable in writing and understanding code, yet it is
rarely formally taught\cite{doi:10.1080/08993400802114581}.  We
typically teach students programming structures, concepts, and
languages, but leave them to learn the tools they use to write code by
themselves.  This approach often works well---a programmer's choice of
tools is often \textit{very} personal and students figure out how to
configure an individualized workflow.  Perhaps because debuggers are
tools, students are often expected to learn them with minimal
guidance.  Unlike editors or reference guides however, effectively
using a debugger requires a set of high-level, platform agnostic,
teachable skills.  Teaching these skills is effective, and translates
into better, faster, debugging and
programming\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}.
Teaching these debugging skills collaboratively will likely offer the
same confidence and correctness benefits realized by teaching
programming collaboratively
\cite{10.1145/1026487.1008043}\cite{10.1145/1145287.1145293}.

@@

The collaborative debugger aims to provide a similar experience to
tools like Replit.  Students should be able to start a debugging
session, easily invite other students, and interact with the debugger
in a way that makes it easy to share their knowledge.  By allowing
multiple users to interact with the same debugging session together in
real-time, the collaborative debugger lets students share experience
and work through problems together.  The debugger also makes it easy
for teachers to demonstrate debugging techniques remotely to multiple
students, any of whom can also interact with the debugger.  Since
debugging sessions are hosted remotely, users can start a session in
one location and resume it later from a different computer.
\par

@@

\begin{enumerate}
\item \lstinline{hash}: an example of classic pointer confusion.  The
  hashtable implementation included with the program is slightly
  bugged, where failing to call \lstinline{strncpy} results in the
  value for each key-value pair in the hash pointing to the same
  location in memory.  This is particularly useful example for learning
  about setting watchpoints and breakpoints, as well as basic
  procedures to step through code.
\item \lstinline{smash}: a brief example of stack smashing.  A badly
  designed call to \lstinline{sscanf} results in a portion of the call
  stack being overwritten, and the process is killed as a result of
  stack smashing.  This example focuses on printing and examining
  memory locations.
\item \lstinline{memoization}: a program that demonstrates the
  difference between a recursive, iterative, and memoized recursive
  implementation of the classic nth Fibonacci number function.  This
  example focuses on examining the stack under conditions produced by
  different types of functions.
\end{enumerate}

@@

The aim of the collaborative debugger is to provide students with an
experience they can directly reference when using similar debuggers in
the future, as well as one which is seamless to encourage the
acquisition of debugging skills.  Creating such a platform will
hopefully encourage both students and teachers to explore the
undertaught area of debugging.

@@

Before diving into the construction of the frontend webapp, it may be
helpful to look at the steps two students would take to debug the
\lstinline{memoization} example program remotely:

@@

Luckily, the speed and size benefits of rr lend themselves well to
non-local execution.  In conjunction with Kubernetes (\ref{k8s}), it
takes a few seconds to create a new container (\ref{docker}) running
rr and connect to web clients.

@@

React makes it easy to create simple webapps with no build
configuration necessary through the use of the
\lstinline{create-react-app} command.  This creates a structure and
toolchain for a basic one-page application---suitable for the
collaborative debugger.  Once this structure was created and any
superfluous files/code were removed, it's wasn't necessary for the
purposes of the collaborative debugger to directly edit any files
except two: \lstinline{App.js} and \lstinline{App.css}.  The
pre-defined build process takes care of transforming the React code
defined in \lstinline{App.js} and the styles defined in
\lstinline{App.css} into portable JavaScript, HTML, and CSS.
\par

@@

As shown in Figure \ref{frontend:rrterm}, the primary piece of the
collaborative debugger's frontend is an interface to the rr debugger
backend.  This interface currently consists of a view of the user's
source code and a terminal interface to rr.  These are contained
within the most important class in the frontend: \lstinline{RRTerm}.
This section will outline the process of sending an rr command to the
backend and displaying a response.
\par

In debugging the \lstinline{memoization} example program, users will
want to display the call stack from within each implementation of the
nth Fibonacci number function.  To stop in \lstinline{fib_rec}, the
recursive implementation, one user will want to set a breakpoint at
line 52 in the source file.  To do this they will type the command
\lstinline{break 52} and hit enter.  The command will be sent to the
backend, and all users in the same debug session (17734 in Figure
\ref{frontend:rrterm}) will see the debuggers response:\\
\lstinline{Breakpoint 1 at 0x561c8fede303: file memo.c, line 52.}\\
Users who did not send the command will see it displayed along with
the response.

@@

Care has been taken to design the frontend in a modular fashion so
that it can be easily extended in the future.  By sending additional
data about program state in the response and creating new components,
visualization tools could be easily added.  If further authentication
features or the ability to edit source code are desired in the future,
relevant components can be updated without the need to change the
entire application.  Since the collaborative debugger depends heavily
on interactions between it's frontend and backend, similar care has
been taken in designing the backend, which will be described in the
following sections.

@@

Kubernetes is the de facto standard in container orchestration
software.  It provides a layer of abstraction on top of normal
containers, like those created by Docker.  By bundling one or more
closely linked containers into a ``Pod'', Kubernetes is able to manage
deployment and re-deployment of applications running inside
containers.  It is trivial to create new Pods, or to create multiple
Pods running the same application as needed within a Kubernetes
cluster \cite{k8s}.  The speed at which even relatively large Pods can
be created and the inherent security provided by containerization
drove the decision to create a new Pod on the fly for each debugging
instance in the collaborative debugger.  This allows the debugger to
provide a similar level of convenience to existing collaborative
tools, such Replit (\ref{exisitingcollab}).
\par

@@

\subsubsection{Cluster Selection and Configuration}

Though Kubernetes aims to be largely platform-agnostic, the
requirements of rr impose some restrictions on cluster setup and
configuration.  Clusters, even those running inside a VM, must be run
on machines using relatively modern Intel x86 CPUs (Nehalem and
beyond).  The clusters must run on an operating system using Linux
kernel version 3.11 or higher \cite{rr-repo}.  Finally, in order for
rr to be able to work efficiently, the
\lstinline{kernel.perf_event_paranoid} parameter must be set to 1
\cite{rr-repo}.  This should be done on every node in the cluster
which will run RR Debug Sessions.  For the purposes of development, it
has been set to 1 on all nodes in the collaborative debugger cluster.

@@

\begin{enumerate}
\item The base image is defined. The RR Message Server and RR Debug
  Session images are based on the latest Ubuntu image.  This is
  particularly necessary for the RR Debug Session image, as rr's
  low-level nature necessitates somewhat frequent updates as changes
  are made to the Linux kernel.  The Frontend/API Server image is
  based on the latest Node image, and the Database image is based on
  the latest MongoDB image.
\item Second, any necessary packages are installed.  For the RR Debug
  Pod image, rr is compiled from source and installed.
\item A non-root user is created if necessary.
\item Program files are copied over and a file structure is created.
  Packages that don't rely on the base images builtin package manager
  are installed at this time.  In the example above, these include
  Python packages.
\item A startup command is defined.
\end{enumerate}

@@

At the heart of the collaborative debugger is the RR Debug Session.
Every time a user wants to debug a new program execution, an RR Debug
Session Pod is dynamically created by the API server.  The new Pod is
assigned a unique five digit identification number when it is created.
This five digit number, is used as the channel for the RR Debug
Session, separating its communication from other RR Debug Session
Pods.  Clients, RR Message Server Pods, and the RR Debug Session Pod
connect through the RR Message Load Balancer.  Clients and RR Debug
Sessions send messages using the channel assigned to the pod at
creation time.  A diagram of one full communication cycle between two
clients and an RR Debug Session Pod is outlined in the diagram below:

@@

  \includegraphics[scale=1]{rr_detailed}
  \centering
  \caption{RR Message Server Communication Cycle}
  \label{rr:detailed}
\end{figure}

@@

\begin{enumerate}
\item Client A, already having connected to the channel that
  corresponds to it's current RR Debug Session (channel 12345 in the
  example), sends an rr command.
\item The RR Message Server receives the command and emits a message
  that RR Debug Session Pods are equipped to receive on the same
  channel.  In practice, since only one RR Debug Session Pod is ever
  on a channel, this equates to emitting a message to the Pod
  directly.
\item The RR Debug Session Pod receives the message, and passes the
  command to its instance of RRInterface.  When it receives a
  response, it emits the output from rr along with other debugging
  information and the channel.
\item The RR Message Server receives the response, and emits a
  response message that clients are equipped to receive on the
  corresponding channel.
\item Clients A and B are connected to the corresponding channel, so
  they both receive the response.
\end{enumerate}

@@

@@

In order to ``support the development of systems which use the
debugger as just one small component of a larger system'', gdb
provides a machine-oriented interface called gdb/mi \cite{gdbman}. rr
supports interaction through gdb/mi, and using the interface was a
natural choice for the collaborative debugger.  In addition to being
far easier to interact with from within a program, the structured,
machine-friendly output of gdb/mi lends itself in particular to future
development of visualization aids in the collaborative debugger.
\par
To parse rr output into Python dictionaries and to easily control rr
as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging Pod.
pygdbmi's abstraction simplifies programmatically controlling rr.  A
Pod can receive a command from the client, pass it to rr, and respond
without having to deal with parsing gdb/mi output or with directly
managing the rr process.

@@

The process for building RR Debug Session example images is differs
slightly from other images used by the collaborative debugger.
Currently, three example RR Debug Session images are available for
use.  Example container images are based on the primary RR Debug
Session image, RR Translation.  The build process for this image,
based on the lasted official Ubuntu container image, installs rr,
python, and all necessary packages as well as the app that will run on
the final image, \lstinline{rrtranslation.py}.

@@

The Dockerfile shown above is for a simple hash program RR Debug
Session example.  The build process copies over any necessary files,
as well as a startup script:

@@

Current example pod startup scripts compile the program to be debugged
and record execution when the pod is created.  It is easy to create a
more specific example to be replayed, if exact reproduction of
syscalls or other non-deterministic behavior is desired.  The
example's creator can record an rr session locally ahead of building
the example Pod and copy the recording into the appropriate folder
during the build process.  Since rr recordings are portable, the
example will replay without issue in the RR Debug Session Pod.

@@

The RR Translation program consists of two main components: a socket.io
client, \lstinline{sio}, and an instance of the
\lstinline{RRInterface} class, \lstinline{rri}, that controls the rr
subprocess and parses interactions.  Simple code to read information
about the current source file currently exists within
\lstinline{sio}'s \lstinline{'rr_command'} event handler, but should
be broken out into its own class if more complexity is added.
\par

The RR Translation program's socket.io client instance begins by
connecting to the RR Message Server.  It immediately emits a
\lstinline{'join_channel'} message, using the channel passed in from
the Kubernetes API call on it's creation.

@@

When \lstinline{write} is called, it determines if the command to be
executed is forbidden.  Currently only one gdb command,
\lstinline{shell}, is disallowed.  Parsing the output from
\lstinline{shell} is difficult, it is of virtually no use since
\textit{recordings}, not currently executing programs that
\lstinline{shell} could effect are being debugged, and it opens the
door to security issues.  Though containerization means that the most
users could probably do is render their own debug session useless, the
downsides of \lstinline{shell} far outweigh the benefits.
\par

@@

Since the frontend currently consists of a terminal interface to rr
and a view of the current source code, write does not need to return
any information from rr other than console output.  Future versions of
the collaborative debugger that support visualization tools could use
gdb/mi commands such as \lstinline{-stack-list-frames} to retrieve
information to be used by frontend visualization tools.
\lstinline{console_output} extracts user-relevant output from the
dictionary returned by \lstinline{GDBController}'s \lstinline{write}
method.  The \lstinline{get_full_rr_response} method ensures that all
relevant output has been received from the rr subprocess before
returning.  Write reduces the lengthy dictionary that would be
produced by a command as simple as \lstinline{break main} into a few
lines of user-relevant output.
\par

@@

Effort is taken in the design of the RR Debug Session and RR Message
Server to ensure students receive a pleasant and near-identical
experience to debugging locally in the future.  The system of
inheritance used to create example debug session pods is designed to
be as simple as possible, though a more fully featured frontend could
simplify this process further in the future.

@@

  \includegraphics[scale=1]{login_request}
  \centering
  \caption{API Request Process---Login}
  \label{rr:detailed}
\end{figure}

@@

All the API request handlers listed above return an error message in
the case of an error occurring during the processing of an API request.
Since most errors that could occur would either be the result of
unanticipated issues on the part of the user, or of some unforeseen
catastrophic internal error, informing users of error specifics would
be unhelpful.  Some \lstinline{TranslationPodManager} and
\lstinline{UserManager} functions throw specific errors in the event
of duplicate usernames, channels that do not exist, etc.  These errors
are caught and meaningful error messages are returned to the frontend
webapp, which can then pass them on to the user.
\par

@@

MongoDB keeps \lstinline{_id}s unique.  In addition, the collaborative
debugger defines uniqueness constrains on \lstinline{users:name},
\lstinline{pods:channel}, and \lstinline{examples:name} when the
database is created.

@@

\subsubsection{Processing a POST Request---Dynamic Pod Creation}

@@

First, \lstinline{DatabaseController}'s \lstinline{add_pod} method is
called to insert a new \lstinline{pods} record into the database.
\lstinline{add_pod} randomly generates a new five digit
\lstinline{channel} for the Pod, using the uniqueness constraint
imposed on \lstinline{pods:channel} to ensure an unused
\lstinline{channel} is generated.  If no channels are open an error is
raised.  Otherwise, the users passed to \lstinline{add_pod} are bound
to the new Pod in the database, and \lstinline{channel} is returned.
This process has the effect of limiting the number of simultaneous RR
Debug Sessions that can be in use at the same time to 89999, and of
slowing as more channels are used.  Given the few current users, the
likelihood of a collision when generating a new channel is low enough
that a more advanced method is unnecessary.
\par

@@

\subsection{Current Functionality}

@@

Student A enters \lstinline{break 52} to set a breakpoint at line 52,
the base case of \lstinline{fib_rec}.  Student B types \lstinline{run}
and then \lstinline{c} to restart the program and continue to the
breakpoint.\pagebreak

>>>>

\begin{abstract}
  \linespread{1} Debugging is rarely formally taught, despite being
  used by programmers every day.  Research indicates that it is
  valuable to teach debugging, and suggests that teaching it
  collaboratively may be maximally effective.  The goal of this
  project is to create a collaborative debugger.  The debugger aims to
  be the ideal platform to teach and learn debugging.  This paper
  briefly reviews relevant literature covering teaching debugging and
  teaching programming collaboratively.  Most of the paper is devoted
  to the design of the collaborative debugger.
\end{abstract}

@@

\begin{itemize}
\item Keith O'Hara, my senior project advisor, for his constant
  thoughtfulness about my project, his advice about the future, and
  for giving me the opportunity to tutor for his classes.
\item Mar Hicks and Kyle Hale, professors whose classes at IIT changed
  my perspective on computer science.
\item All my professors at Bard, particularly Olga Voronina, Wakako
  Suzuki, and Nathan Shockey, whose encouragement of my interests
  outside of computer science has been instrumental in continuing my
  passion for it.
\item My family and friends, for their constant love and support.
\end{itemize}

@@

Debugging is invaluable in writing and understanding code, yet it is
rarely formally taught\cite{doi:10.1080/08993400802114581}.  Students
are typically taught programming structures, concepts, and languages,
but are left to learn the tools they use to write code alone.  This
approach often works well---a programmer's choice of tools is often
\textit{very} personal and students figure out how to configure an
individualized workflow.  Perhaps because debuggers are tools,
students are often expected to learn them with minimal guidance.
Unlike editors or reference guides however, effectively using a
debugger requires a set of high-level, platform agnostic, teachable
skills.  Teaching these skills is effective, and translates into
better and faster debugging and
programming\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}.
Teaching these debugging skills collaboratively will likely offer the
same confidence and correctness benefits realized by teaching
programming collaboratively
\cite{10.1145/1026487.1008043}\cite{10.1145/1145287.1145293}.

@@

The collaborative debugger aims to provide a similar experience to
tools like Replit.  Students should be able to start a debugging
session, easily invite other students, and interact with the debugger
in a way that makes it easy to share their knowledge.  By allowing
multiple users to interact with the same debugging session together in
real-time, the collaborative debugger lets students share experience
and work through problems together.  The debugger also makes it easy
for teachers to demonstrate debugging techniques remotely to multiple
students, any of whom can also interact with the debugger.  Since
debugging sessions are hosted remotely, users can start a session in
one location and resume it later from a different computer.  This is
particularly beneficial for students who may not always have access to
a computer or a machine capable of running the programs they wish to
debug.
\par

@@

\begin{enumerate}
\item \lstinline{hash}: an example of classic pointer confusion.  The
  hashtable implementation included with the program is slightly
  bugged, where failing to call \lstinline{strncpy} results in the
  value for each key-value pair in the hash pointing to the same
  location in memory.  This is a particularly useful example for
  learning about setting watchpoints and breakpoints, as well as basic
  procedures to step through code.
\item \lstinline{smash}: a brief example of stack smashing.  A badly
  designed call to \lstinline{sscanf} results in a portion of the call
  stack being overwritten, and the process is killed as a result of
  stack smashing.  This example focuses on printing and examining
  memory locations.
\item \lstinline{memoization}: a program that demonstrates the
  difference between a recursive, iterative, and memoized-recursive
  implementation of the classic nth Fibonacci number function.  This
  example focuses on examining the stack under conditions produced by
  different types of functions.
\end{enumerate}

@@

The aim of the collaborative debugger is to provide students with an
experience they can directly reference when using similar debuggers in
the future, as well as one which is seamless to encourage the
acquisition of general debugging skills.  Creating such a platform
will hopefully encourage both students and teachers to explore the
undertaught area of debugging.

@@

Before diving into the construction of the frontend webapp, it may be
helpful to look at the steps two students would take to debug the
\lstinline{memoization} example program remotely.  This example is
covered in more detail in \ref{current}.

@@

Luckily, the speed and size benefits of rr lend themselves well to
non-local execution.  In conjunction with tools used to create the
backend of the collaborative debugger it takes a few seconds to create
a new debug session running rr and connect it to the frontend webapp
discussed in this section.

@@

% The frontend webapp communicates with a distributed backend.  A simple
% diagram of the entire system is below.

% The collaborative debugger consists of:

% \begin{enumerate}
% \item A frontend web app that presents a debugging interface to the
%   end user.
% \item A distributed backend split into three parts:
%   \begin{enumerate}
%   \item 
%   \item Pods running a frontend server written in Node.js which works
%     in tandem with an API server created using Flask
%     (\ref{flask/node}).  The API server manages creation and
%     destruction of debug sessions, as well as authentication.
%   \item A MongoDB (\ref{mongodb}) database.
%   \end{enumerate}
% \end{enumerate}

React makes it easy to create simple webapps with no build
configuration necessary through the use of the
\lstinline{create-react-app} command.  This creates a structure and
toolchain for a basic one-page application---suitable for the
collaborative debugger.  Once this structure was created and any
superfluous files/code were removed, it wasn't necessary for the
purposes of the collaborative debugger to directly edit any files
except two: \lstinline{App.js} and \lstinline{App.css}.  The
pre-defined build process takes care of transforming the React code
defined in \lstinline{App.js} and the styles defined in
\lstinline{App.css} into portable JavaScript, HTML, and CSS.
\par

@@

As shown in Figure \ref{frontend:rrterm}, the primary piece of the
collaborative debugger's frontend is an interface to the rr debugger
backend.  This interface currently consists of a view of the user's
source code and a terminal interface to rr.  These are contained
within the most important class in the frontend: \lstinline{RRTerm}.
This section will outline the process of sending an rr command to the
backend and displaying a response.  Apart from sending simple API
requests to start sessions or log users in and out, this communication
process is where the frontend webapp spends most of its time.
\par

In debugging the \lstinline{memoization} example program, users will
want to display the call stack from within each implementation of the
nth Fibonacci number function.  To stop in \lstinline{fib_rec}, the
recursive implementation, one user will want to set a breakpoint at
line 52 in the source file.  To do this they will type the command
\lstinline{break 52} and hit enter.  The command will be sent to the
backend, and all users in the same debug session (17734 in Figure
\ref{frontend:rrterm}) will see the debuggers response:\\
\lstinline{Breakpoint 1 at 0x561c8fede303: file memo.c, line 52.}\\
Users who did not send the command \lstinline{c} will see it displayed
along with the response.

@@

Care has been taken to design the frontend in a modular fashion so
that it can be easily extended in the future.  Visualization tools
could be easily added by sending additional data about program state
in the response and creating new components.  If further
authentication features or the ability to edit source code are desired
in the future, relevant components can be updated without the need to
change the entire application.  Since the collaborative debugger
depends heavily on interactions between its frontend and backend,
similar care has been taken in designing the backend, which will be
described in the following sections.

@@

Kubernetes is the de facto standard in container orchestration
software.  It provides a layer of abstraction on top of normal
containers, like those created by Docker.  By bundling one or more
closely linked containers into a ``Pod'', Kubernetes is able to manage
deployment and re-deployment of applications running inside
containers.  It is trivial to create new Pods (or multiple Pods
running the same application) as needed within a Kubernetes cluster
\cite{k8s}.  The speed at which even relatively large Pods can be
created and the inherent security provided by containerization drove
the decision to create a new Pod on the fly for each debugging
instance in the collaborative debugger.  This allows the debugger to
provide a similar level of convenience to existing collaborative
tools, such Replit (\ref{exisitingcollab}).
\par

@@

\subsubsection{Cluster Requirements and Configuration}

Though Kubernetes aims to be largely platform-agnostic, the
requirements of rr impose some restrictions on cluster setup and
configuration.  Clusters, even those running inside a VM, must be run
on machines using relatively modern Intel x86 CPUs (Nehalem and
beyond).  The clusters must run on an operating system using Linux
kernel version 3.11 or higher \cite{rr-repo}.  Finally, in order for
rr to be able to work efficiently, the
\lstinline{kernel.perf_event_paranoid} parameter must be set to
\lstinline{1} \cite{rr-repo}.  This should be done on every node in
the cluster which will run RR Debug Sessions.  For the purposes of
development, it has been set to \lstinline{1} on all nodes in the
collaborative debugger cluster.

@@

\begin{enumerate}
\item The base image is defined. The RR Message Server and RR Debug
  Session images are based on the latest Ubuntu image.  This is
  particularly necessary for the RR Debug Session image, as rr's
  low-level nature necessitates somewhat frequent updates as changes
  are made to the Linux kernel.  The Frontend/API Server image is
  based on the latest Node image, and the Database image is based on
  the latest MongoDB image.
\item Second, any necessary packages are installed.  For the RR Debug
  Session image, rr is compiled from source and installed.
\item A non-root user is created if necessary.
\item Program files are copied over and a file structure is created.
  Packages that don't rely on the base image's built in package
  manager are installed at this time.  In the example above, these
  include Python packages.
\item A startup command is defined.
\end{enumerate}

@@

At the heart of the collaborative debugger is the RR Debug Session.
Every time a user wants to debug a new program execution, an RR Debug
Session Pod is dynamically created by the API server.  The new Pod is
assigned a unique five digit identification number when it is created.
This five digit number is used as the channel for the RR Debug
Session, separating its communication from other RR Debug Session
Pods.  Clients, RR Message Server Pods, and the RR Debug Session Pod
connect through the RR Message Load Balancer.  Clients and RR Debug
Sessions send messages using the channel assigned to the pod at
creation time.  A diagram of one full communication cycle between two
clients and an RR Debug Session Pod is outlined in the diagram below:

@@

  \includegraphics[scale=.8]{rr_detailed}
  \centering
  \caption{RR Message Server Communication Cycle}
  \label{rr:detailed}
\end{figure}

@@

\begin{enumerate}
\item Client A, already having connected to the channel that
  corresponds to it's current RR Debug Session (channel 12345 in the
  example), sends an rr command.
\item The RR Message Server receives the command and emits a message
  that RR Debug Session Pods are equipped to receive on the same
  channel.  In practice, since only one RR Debug Session Pod is ever
  on a channel, this equates to emitting a message to the Pod
  directly.
\item The RR Debug Session Pod receives the message, and passes the
  command to its instance of RRInterface, the class which controls rr.
  When it receives a response, it emits the output from rr along with
  other debugging information and the channel.
\item The RR Message Server receives the response, and emits a
  response message that clients are equipped to receive on the
  corresponding channel.
\item Clients A and B are connected to the corresponding channel, so
  they both receive the response.
\end{enumerate}

@@

RR Debug Sessions primarily manage and communicate with rr, which has
already been covered in detail in \ref{rr}.  pygdbmi is used for this
process.

@@

In order to ``support the development of systems which use the
debugger as just one small component of a larger system'', gdb
provides a machine-oriented interface called gdb/mi \cite{gdbman}. rr
supports interaction through gdb/mi, and using the interface was a
natural choice for the collaborative debugger.  In addition to being
far easier to interact with from within a program, the structured,
machine-friendly output of gdb/mi lends itself in particular to future
development of visualization aids in the collaborative debugger.
\par
To parse rr output into Python dictionaries and to easily control rr
as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging Pod.
pygdbmi's abstraction simplifies programmatically controlling rr.  A
Pod can receive a command from the client, pass it to rr, and respond
without having handle directly parsing gdb/mi output or managing the
rr process.

@@

The process for building RR Debug Session example images is differs
slightly from other images used by the collaborative debugger.
Currently, three example RR Debug Session images are available for
use: \lstinline{hash}, \lstinline{smash}, and \lstinline{memoization}.
Example container images are based on the primary RR Debug Session
image, RR Translation.  The build process for this image, based on the
lasted official Ubuntu container image, installs rr, python, and all
necessary packages as well as the app that will run on the final
image, \lstinline{rrtranslation.py}.

@@

The Dockerfile shown above is for the \lstinline{hash} RR Debug
Session example.  The build process copies over any necessary files,
as well as a startup script:

@@

Current example pod startup scripts compile the program to be debugged
and record execution when the pod is created.

@@

The RR Translation program consists of two main components: a
Socket.IO client, \lstinline{sio}, and an instance of the
\lstinline{RRInterface} class, \lstinline{rri}, that controls the rr
subprocess and parses interactions.  Simple code to read information
about the current source file currently exists within
\lstinline{sio}'s \lstinline{'rr_command'} event handler, but should
be broken out into its own class if more complexity is added.
\par

The RR Translation program's socket.io client instance begins by
connecting to the RR Message Server.  It immediately emits a
\lstinline{'join_channel'} message, using the channel number passed in
by the Kubernetes API on Pod creation (\ref{dynamicc}).

@@

When \lstinline{write} is called, it determines if the command to be
executed is forbidden.  Currently only one gdb command,
\lstinline{shell}, is disallowed.  Parsing the output from
\lstinline{shell} is difficult, it is of virtually no use since
\textit{recordings}, not currently executing programs that
\lstinline{shell} could effect, are being debugged.  It also opens the
door to security issues.  Though containerization means that the most
users could probably do is render their own debug session useless, the
downsides of \lstinline{shell} far outweigh the benefits.
\par

@@

Since the frontend currently consists of a terminal interface to rr
and a view of the current source code, \lstinline{write} does not need
to return any information from rr other than console output.  Future
versions of the collaborative debugger that support visualization
tools could use gdb/mi commands such as \lstinline{-stack-list-frames}
to retrieve information to be used by frontend visualization tools.
\lstinline{console_output} extracts user-relevant output from the
dictionary returned by \lstinline{GDBController}'s \lstinline{write}
method.  The \lstinline{get_full_rr_response} method ensures that all
relevant output has been received from the rr subprocess before
returning.  \lstinline{write} reduces the lengthy dictionary that
would be produced by a command as simple as \lstinline{break main}
into a few lines of user-relevant output.
\par

@@

Effort is taken in the design of the RR Debug Session and RR Message
Server to ensure students receive a pleasant and near-identical
experience to debugging locally in the future.  The system of
inheritance used to create example debug session pods is designed to
be as simple as possible, though a more fully featured frontend could
streamline this process further in the future.

@@

  \includegraphics[scale=.9]{login_request}
  \centering
  \caption{API Request Process---Login}
  \label{rr:detailed}
\end{figure}

@@

All the API request handlers listed above return an error message in
the case of an error occurring during the processing of an API
request.  Since most errors that could occur would either be the
result of unanticipated issues on the part of the user, or of some
unforeseen catastrophic internal error, informing users of error
specifics is often unhelpful.  Some \lstinline{TranslationPodManager}
and \lstinline{UserManager} functions throw specific errors in the
event of duplicate usernames, channels that do not exist, etc.  These
errors are caught and meaningful error messages are returned to the
frontend webapp, which can then pass them on to the user.
\par

@@

MongoDB keeps \lstinline{_id}s unique.  In addition, the collaborative
debugger defines uniqueness constrains on \lstinline{users.name},
\lstinline{pods.channel}, and \lstinline{examples.name} when the
database is created.

@@

\subsubsection{Processing a POST Request---Dynamic Pod Creation} \label{dynamicc}

@@

First, \lstinline{DatabaseController}'s \lstinline{add_pod} method is
called to insert a new \lstinline{pods} record into the database.
\lstinline{add_pod} randomly generates a new five digit
\lstinline{channel} for the Pod, using the uniqueness constraint
imposed on \lstinline{pods.channel} to ensure an unused
\lstinline{channel} is generated.  If no channels are open an error is
raised.  Otherwise, the users passed to \lstinline{add_pod} are bound
to the new Pod in the database, and \lstinline{channel} is returned.
This process has the effect of limiting the number of simultaneous RR
Debug Sessions that can be in use at the same time to 89,999, and of
slowing as more channels are used.  Given the few current users, the
likelihood of a collision when generating a new channel is low enough
that a more advanced method is unnecessary.
\par

@@

\subsection{Current Functionality} \label{current}

@@

Student A enters \lstinline{break 52} to set a breakpoint at line 52,
the base case of \lstinline{fib_rec}.  Student B types \lstinline{run}
and then \lstinline{c} to restart the recording and continue to the
breakpoint.\pagebreak
